\subsection{Linear programs}

    Let $A\in\reals^{n\times k}$, $b\in\reals^{n}, c\in\reals^{k}$. A linear program is a constrained optimisation problem of the following form
    \footnote{This form of a linear program is called standard form. }
    \begin{align*}
        (LP)
        \begin{cases}
            \min_{x\in\reals^{k}}cx^{T} ~\text{ subject to}\\
            ~~~~ Ax^{T}\leq b^{T}\\
            ~~~~ x\geq \mathbf{0}
        \end{cases}
    \end{align*}
    where $\mathbf{0}$ denotes the vector containing only zeros and all the inequality relations are to be understood componentwise. The polyhedron
    \begin{equation*}
        \mathcal{F} = \{x\in\reals^{k} : x\geq\mathbf{0}, Ax^{T}\leq b^{T}\}
    \end{equation*}
    is called the feasible set of $(LP)$. It can be shown that any solution to $(LP)$, if it exists, can be found among the vertices of $\mathcal{F}$. For small
    problems this result leads in a feasible algorithm to solve $(LP)$:
    \begin{enumerate}
        \item Find all the vertices of $\mathcal{F}$
        \item Evaluate the objective function $f(x) = cx^{T}$ among the vertices
        \item The solution of $(LP)$ is the vertex minimising $f$.
    \end{enumerate}
    For large problems however the numerical costs of this approach are too large. This is 
    why the simplex method comes into play, which tries to reduce the number of vertices which have to be visited.

    \begin{example}
        Consider the optimisation problem
        \begin{align*}
            \begin{cases}
                \min -x_{1} - 3x_{2} ~\text{ subject to}\\
                ~~~~ 2x_{1} + 3x_{2}\leq 6\\
                ~~~~ x_{1} - x_{2}\geq -1\\
                ~~~~ x_{1}, x_{2} \geq 0
            \end{cases}                        
        \end{align*}
        In order to write this problem as linear program let us first note that
        \begin{equation*}
            x_{1} - x_{2}\geq -1 \iff -x_{1} + x_{2}\leq 1.
        \end{equation*}
        Definining the matrix $A$, and the vectors $b, c$ via
        \begin{align*}
            A = 
            \begin{bmatrix}
                2 & 3\\
                -1 & 1\\
            \end{bmatrix}, 
            ~b = (6, 1), ~c = (-1, -3)
        \end{align*}
        we infer that the given problem can be rewritten as linear program. The feasible set is depicted in the upfollowing figure. \newline
        \begin{tikzpicture}
            \begin{axis}
            [
                xmin = -1, xmax = 4,
                ymin = -1, ymax = 3,
                xlabel = {$x_{1}$},
                ylabel = {$x_{2}$},
                grid = both,
                grid style = {line width = .1pt, draw = darkgray!10},
                major grid style = {line width = .2pt, draw = darkgray!50},
                axis lines = middle,
                minor tick num = 4,
                % enlargelimits = {abs = 0.5},
                samples = 100,
                % domain = -20:20,
            ]
            \filldraw[blue, pattern = north west lines, pattern color = blue] (0, 0) -- (0, 1) -- (0.6, 1.6) -- (3, 0) -- cycle;
            \end{axis}
        \end{tikzpicture}
        \newline
        As the figure shows, the feasible set has four vertices, namely $(0, 0), (1, 0), (3 / 5, 8 / 5), (3, 0)$. Moreover we infer that the 
        feasible set is compact. Thus thelinear problem has a solution, which can be found among the vertices of its feasible set. By comparing the 
        values of the objective function $f(x_{1}, x_{2}) = -x_{1} - 3x_{2}$ at the vertices we find that $(3 / 5, 8 / 5)$ solvers the problem.
    \end{example}
